An Adolescent Mini-Neptune in the Kepler Field

### 2021-08-12 paper/ms.tex
update ms and CMD age

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.
  To our knowledge, Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The Kepler
  photometry seems to show two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are well-poised to expand the census of
  age-dated planets, and our understanding of planetary evolution.
\end{abstract}

% While thousands of exoplanets have been discovered orbiting nearby
% stars, the vast majority of them are several billion years old.  This
% makes it difficult to test theories for the origins of the different
% families of planets, since many evolutionary processes are expected to
% operate on timescales of less than 100 million years.  The K2 and TESS
% missions have enabled advances on this front [1,2,3,4,5,6].  The
% Kepler mission however has not yielded any planets with precise ages
% below one gigayear [7,8].  Here, we combine kinematic data from Gaia
% with stellar rotation periods from TESS to show that Kepler 1627 is a
% member of the 35 ± 10 million year old δ Lyr cluster.  A
% $3.6\,R_\oplus$ mini-Neptune transits the G8V primary on a 7.2 day
% orbit, and it is therefore the youngest planet yet found by the main
% Kepler mission.  The precision of the Kepler photometry lets us detect
% two anomalies.  First, unresolved starspot-crossing events distort the
% average transit shape, and induce transit timing variations that imply
% the planet is likely on a prograde orbit [9].  Second, although the
% flare arrival time distribution is most likely Poissonian, there may
% be hints of an excess of flares at the orbital and synodic periods.
% The δ Lyr cluster is one of many stellar groups whose nature has been
% clarified using the Gaia data; many other new clusters are expected to
% overlap datasets of confirmed and candidate planets.  Further studies
% that combine Gaia kinematics and TESS rotation periods are therefore
% well-poised to expand the census of age-dated planets, and our
% understanding of planetary evolution.

@@

%TODO: might want to revise this text to better frame the boil-off phase...
%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of
molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ near the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets, measuring their masses, and detecting their
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster
(Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
Figure~\ref{fig:skychart} shows the overlap between the Kepler field
and the cluster.

A previous clustering analysis of the Gaia data reported
that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member
of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
validation of its close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative
host cluster most closely (Section~\ref{sec:cluster}).  We find that the $\delta$
Lyr cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself
(Section~\ref{sec:planet}), we confirm that despite a previously
unreported binary companion, it is indeed a mini-Neptune orbiting the
G-dwarf primary star.  The presence of a correlation
between transit timing variations and the local light curve slope
renders any other explanation implausible.  We conclude by
highlighting some broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f15a.pdf}
			\includegraphics[width=0.469\textwidth]{f15b.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} Color-magnitude diagram of 
    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
    The right panel provides a closer view of stars on the
    pre-main-sequence.
    The \cn\ and IC\,2602 are approximately the same isochronal age.
    {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
    colors, with the Pleiades (125\,Myr)
    and Praesepe (650\,Myr)
    shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.
    Kepler\,1627 is shown with a yellow star.
     {\bf todo: caption improve.}.
   \label{fig:age}
	}
\end{figure*}

% paragraph 1: HR 
To assess the age of the cluster and our star of interest,
we used a combination of the isochronal, gyrochronal, and
lithium-based age-dating.

@@

%FIXME FIXME TODO TODO get the ages measured

The immediate impression from Figure~\ref{fig:age}
is that the \cn\ and IC\,2602 overlap, and therefore are approximately the same isochronal
age.  In our exploration, we also compared against
$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and the
Upper-Centaurus-Lupus component of the Sco OB2 association (UCL;
$\approx10$-$15$\,Myr; \citealt{damiani_stellar_2019}).  
The pre-main-sequence M dwarfs of the \cn\ were intermediate
between these latter two clusters.  To quantify this into an age measurement
with an uncertainty, we adopted the empirical procedure
developed by \citet{gagne_mutau_2020} in their Section~6.  In brief,
we fitted the pre-main-sequence loci of a set of
reference clusters, and the locus of the target \cn\ was
then modeled as a piecewise linear combination of these reference
clusters.  For our reference clusters, we adopted members of UCL, IC\,2602, and the
Pleaides from \citet{damiani_stellar_2019} and
\cite{CantatGaudin2018a} respectively.
We removed binaries by requiring
$\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$ below the
80$^{\rm th}$ percentile of each cluster's distribution, and by
manually excluding stars that were obvious photometric binaries in the
CMD.  We then passed a moving box average and standard deviation
across the CMD in 0.10\,mag bins, fitted a univariate spline to the
binned values, and assembled a piecewise grid of hybrid isochrones
spanning the ages between UCL to the Pleiades following Equations~6
and~7 from \citet{gagne_mutau_2020}.

The ages returned by this procedure depend on the ages assumed for
each reference cluster.
In our case, the IC\,2602 cluster is most important.
As noted by \citet{bouma_cluster_2020}, plausible ages for IC\,2602
span 30\,Myr to 46\,Myr, with slightly older ages being preferred by the
lithium-depletion-boundary (LDB) measurements and younger ages by the
main-sequence turn-off
\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
$31^{+5}_{-4}$\,Myr old.
For the converse case of 46\,Myr, the \cn\ would be
$44^{+8}_{-7}$\,Myr old.
We adopt an intermediate 35\,Myr age as our reference, 
yielding an age for the \cn\ of
$35^{+6}_{-5}$\, Myr, though it should be understood based on the
systematic uncertainties that an age of up to $\approx50$\,Myr remains
plausible.  Detailed analyses of the LDB or main-sequence
turn-off in the \cn\ could help clarify the issue, but both are beyond
our scope.

>>>>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining kinematic
  information from Gaia, stellar rotation periods from TESS, and Li
  abundances from ground-based spectra, we show that Kepler\,1627 is a
  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.  To our
  knowledge, Kepler\,1627Ab is the youngest planet with a precise age
  yet found by the main Kepler mission.  From both unresolved
  starspot-crossing events in the Kepler photometry and from a
  ground-based Keck/HIRES transit we find that the planet is likely on
  an orbit nearly aligned with the stellar spin axis.  We also find
  from the short-cadence Kepler photometry that although the flare
  arrival time distribution is probably Poissonian, there may be an
  excess of flares separated by the orbital and synodic periods, and
  linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand
  the census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

%TODO: might want to revise this text to better frame the boil-off phase...
%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of molten
rocky cores \citep{kite_atmosphere_2020} and extended atmospheric
envelopes of hydrogen and helium, are expected to shrink in size by
factors of several over their first $10^8$ years
\citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ near the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
is expected to occur within the first few million years of disk
dispersal \citep{owen_atmospheres_2016}, the combination of stellar
irradiation and internal heat may also power a more
gradual outflow, which can eventually deplete or entirely strip the
envelope from the rocky core
\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  Discovering young
planets, measuring their masses, and detecting their atmospheric
outflows are key steps toward testing this paradigm, which is often
invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by
J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster (Stephenson-1; Theia~73;
\citealt{stephenson_possible_1959}).  Figure~\ref{fig:skychart} shows
the overlap between the Kepler field and the cluster.

A previous clustering analysis of the Gaia data reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
\cn\ \citep{kounkel_untangling_2019}.  Given the statistical
validation of its close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative host cluster more
closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself
(Section~\ref{sec:planet}), we confirm through two independent
approaches that despite a previously unreported binary companion, it
is a mini-Neptune orbiting the G-dwarf primary star.  We conclude by
highlighting some broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f15a.pdf}
			\includegraphics[width=0.469\textwidth]{f15b.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35^{+6}_{-5}$\,Myr old.}  {\it Top:}
    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
    field (background contours).  The right panel provides a closer
    view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
    approximately the same isochronal age.  {\it Bottom:} TESS and
    Kepler stellar rotation period versus dereddened Gaia color, with
    the Pleiades (115\,Myr) and Praesepe (650\,Myr) shown for
    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
    Kepler\,1627 is shown with a yellow star.
   \label{fig:age}
	}
\end{figure*}

To assess the age of the cluster and our star of interest, we used a
combination of the isochronal, gyrochronal, and lithium-based
age-dating.

@@

Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and are
therefore approximately the same age.  In our exploration, we also
compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
association ($\approx$-$16$\,Myr; \citealt{pecaut_star_2016}).  The
pre-main-sequence M dwarfs of the \cn\ were intermediate between these
latter two groups.  To turn this heuristic interpolation into an
age measurement with an uncertainty, we used the empirical procedure
developed by \citet{gagne_mutau_2020}.  In brief, we fitted the
pre-main-sequence loci of a set of reference clusters, and the locus
of the target \cn\ was then modeled as a piecewise linear combination
of these reference clusters.  For our reference clusters, we adopted
members of UCL, IC\,2602, and the Pleaides from
\citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}
respectively.  We removed binaries by requiring $\texttt{RUWE}<1.3$,
$\texttt{radial\_velocity\_error}$ below the 80$^{\rm th}$ percentile
of each cluster's distribution, and by manually excluding stars that
were obvious photometric binaries in the CMD.  We then passed a moving
box average and standard deviation across the CMD in 0.10\,mag bins,
fitted a univariate spline to the binned values, and assembled a
piecewise grid of hybrid isochrones spanning the ages between UCL to
the Pleiades following Equations~6 and~7 from
\citet{gagne_mutau_2020}.

The ages returned by this procedure depend on the ages assumed for
each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.
The age of the IC\,2602 cluster however is the most important
ingredient, since it receives the most weight in the interpolation.
As noted by \citet{bouma_cluster_2020}, plausible ages for IC\,2602
span 30\,Myr to 46\,Myr, with slightly older ages being preferred by
the lithium-depletion-boundary (LDB) measurements and younger ages by
the main-sequence turn-off
\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
be $31^{+5}_{-4}$\,Myr old.  For the converse case of 46\,Myr, the
\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
35\,Myr age as our reference, yielding an age for the \cn\ of
$35^{+6}_{-5}$\, Myr, though it should be understood based on the
systematic uncertainties that an age of up to $\approx50$\,Myr remains
plausible.  Detailed analyses of the LDB or main-sequence turn-off in
the \cn\ could help clarify the issue, but both are beyond our scope.
